Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-06-27 Latex_files/Thesis/chapters/chapter 2/main.tex
27/06/2025 chapters 1,3,4 and structural changes

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two somewhat unconnected
fields; crystallisation theory, which is covered in the 
previous introduction section, and tweezing theory. The 
following chapter summarises the working principles behind 
optical tweezing, the differences in scattering theories, 
and the broader theoretical calculations used throughout the 
rest of the thesis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Electromagnetism and optical tweezing}

Proper understanding of optical tweezing requires a description
of how the trapped particle interacts with the laser. 
From an electromagnetism perspective the laser creates a 
spatially and temporally coherent electric field that scatters 
light when passing though a medium. The laws governing electric 
and magnetic fields are summarised most succinctly via the 
Maxwell equations. The differential forms of which are given 
below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \label{eq:max_start}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}
  \label{eq:max_end}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the 
charge current, and $\epsilon_0$ is the permittivity of free 
space.

These four equations describe how the electric and magnetic fields 
behave and relate to one another at a microscopic level. Any 
discussion of optical trapping is underpinned by the fact that in 
all cases the Maxwell equations must be satisfied at the macroscopic 
level, where one must account for the medium's interactions with the 
EM field. Macroscopic forms of the Maxwell equations are given below
\cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and 
current caused by the presence of free charges in the medium, 
and $\mathbf{D}$ and $\mathbf{H}$ are the displacement and 
magnetising fields respectively. The latter two are defined 
by:
\begin{align}
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

Where $\mu_0$ is the vacuum permeability $\mathbf{P}$ 
is the polarization field and $\mathbf{M}$ is the 
magnetisation field, these fields arise due to the 
bound charges throughout the medium interacting with 
the EM field.

@@

The Lorenz-Mie theory provides an exact solution to the Maxwell 
equations for the scattering caused by an isotropic sphere. This 
theory describes the scattered wave given off by a dielectric 
sphere when incident by a plane wave as a summation of partial 
spherical waves. For any wave equation the vector fields must 
solve the Helmholtz equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a approximate Gaussian beam 
interacting with a sphere of radius $a$, the incident, internal, 
and scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and ${\bf M}_
{nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
spherical harmonics respectively. The superscript denote whether 
the field is inward [(1)] or outward propagating [(3)]. By linearity 
of Maxwell's equations, its possible to relate the expansion 
coefficients of the scattered field directly to those of the incident 
field. As such we can write the scattering coefficients in terms of 
their incident components and the target's 'T-matrix'.
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} & T_{12} \\
		T_{21} & T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where the matrix \textbf{T} accounts for the targets shape, 
and refractive index. The $T$-matrix method can be easily 
applied to any solution of Maxwell's equations, for example 
Eq.~\eqref{eq:scattered} can be instead be rewritten as:
\begin{align}
{\bf E}_{\rm scat}({\bf r})
&=
E_0 \sum_{n}^\infty \sum_{m=-n}^n 
\begin{pmatrix}
	T_{11} & T_{12} \\
	T_{21} & T_{22}
\end{pmatrix}
\begin{pmatrix}
	a_{mn} \\
	b_{mn}
\end{pmatrix}
\begin{pmatrix}
	{\bf N}_{mn}^{(3)} & {\bf M}_{mn}^{(3)}
\end{pmatrix}
\end{align}

In order to compute the optical force and torque applied to a 
given object one simply needs to know the momentum carried by 
the incident and scattered fields. So long as we know the 
expansion coefficients of the incident beam we can find the 
scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
compute the force. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton 
\cite{Crichton2000THEMD}. As an example the time averaged linear 
momentum carried by any incident beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

@@

\noindent
Where $a_{nm}, b_{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size.
Though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are easier to approximate by alternative theories. When the 
particle radius is $\gg \lambda$ the optical force is best described 
by the Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rayleigh Scattering}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered 
field produced by a particle. While one can compute this by 
directly applying Lorenz-Mie theory, describing the interactions 
between focused beams and complicated shapes can be computationally 
slow and memory intensive. Several alternative scattering approaches 
have been developed that allow for faster computation times 
simplifying the scattering problem. While all of these methods will
produce similar results the key difference is in how one represents
the scatterer.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (EBCM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the EBCM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix. Because the T-matrix is based only 
on the properties of the scatterer it can be reused in 
calculations regardless of the incident field in question.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker 
\cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh approximation, though in order to compute 
the optical force one now requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target. 
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Describing the trajectory of a trapped particle}
While describing the optical forces is sufficient for 
theoretical characterisation of an optical trap, in an 
experimental setting the observed motion is not as 
simple. As such it is often far simpler to 
characterise the optical trap by recording the particle's
trajectory (see \ref{sec:position_detection} for 
specifics regarding the equipment), and then fitting it 
to our presumed description of the beam. The trajectory 
is easily understood if we understand the particle's 
diffusive behaviour.

@@

\newpage
\section{Calibration Techniques}
\label{sec:calibration_techniques}
There are several approaches for calibrating and characterizing 
the optical trap, each approach has its drawbacks and benefits 
so each option should be chosen based on what elements want to 
be characterized. The basis for each of these methods stems from 
the analytical solution of the Langevin equation:
\begin{align}
	\label{eq:anylitical_lang}
	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of spherical aggregates}
\label{sec:simulations}
Later chapters cover the dynamics of spherical aggregates and anisotropic 
scatterers, these subjects are particularly difficult to characterise 
using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
As an example consider a symmetric dimer as a paradigmatic aggregate; 
if we consider the Langevin equation for such a aggregate within an optical 
trap we have:
\begin{align}
	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
	\sqrt{2\vec{D}_x}\eta(t)
\end{align}

Vigilante et al compiled together a python package that combines 
both \textit{ott} and \textit{mstm} to simulate the behaviour of 
spherical aggregates within a predefined optical trap 
\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
modules that can be used to model the scattering from different 
shaped particles; however, for aggregates of spheres the resultant 
force is nonsensical (see Chapter 4 for further details). As such 
we rely on the \textit{mstm} for simulating dimers.

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
to perform a systematic study of the dynamics demonstrated by 
asymmetric dimers in both plane and circularly polarised light. 
The simulations have two primary frames of reference that require 
explanation. Firstly, there is the laboratory frame which 
describes the coordinate system, we define the origin of our 
simulation as the focus of our trapping beam. While the choice in 
direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
specifically chosen so that it is aligned with the direction of 
beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation 
state of the particle, orientations are reported as quaternions which 
can be readily transformed into rotation matrices. The matrix describes 
the direction of the particles 3 orientation axis $U_x$, $U_y$, and 
$U_z$. The latter is oriented so that it is collinear with the centres
of the two spheres; the direction of vector $U_z$ is such that it goes 
from the centre of $a_{II}$ to the centre of $a_{I}$. The origin of the 
3 axis' is set on the particle's centre of mass, this is also used to 
define it's displacement from the origin of the laboratory frame. The 
diffusion tensor of the dimer is more complex than a sphere's as we must consider the rotational effects experienced by the dimer. We use the 
analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided 
the solution coefficients depending on the dimer's size ratio 
($a_I/a_{II}$). Rather than recalculate the coefficients for every possible configuration of dimer we use a spline fitting function to approximate 
the diffusion tensor. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulated Quadrant Photodiode}
\label{sec:simulated_QPD}

Beyond merely simulating the forces experienced by a spherical 
aggregate we also developed a simulative quadrant photo diode 
to replicate the results from a typical calibration test. This 
builds upon the work from \cite{Rohrbach2002} which applied 
Lorenz-Mie theory to replicate the response signal of a QPD 
being used in back focal-plane interferometry. Rather than be 
constrained to individual spheres we can now consider the 
expected response from any type of spherical aggregate. The
goal of such simulation software would be to simulate the 
expected experimental results from a complex particle whose 
exact shape and size may be difficult to determine.

In order to simulate a typical experimental set up with a QPD 
installed as a position detection system we need to evaluate 
the total magnitude of the electric field incident on the 
photo-diode surface. While trapping a micro-particle, the 
scattered and incident fields combine together and interfere 
with one another. These fields are collected by a condenser 
lens in the far field limit and are focused onto the QPD 
surface, the total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(\theta_{NA}-\theta(x,y))
\end{align}

The last term is simply a representative step term that defines 
the outer limit by which we evaluate the electric field, this is 
analogous to our condenser lens removing noise from other light 
sources by only accepting light at a specific acceptance angle 
defined by its numerical aperture $NA_c$. Depending on the 
relative size of our particle we can adjust the acceptance angle, 
this has very little effect on the transverse signals, but for 
axial evaluations of a trapped particle the numerical aperture 
should be tuned so that the resultant response curve has negative 
slope in order to allow for axial position detection, the method 
for finding this angle $\theta_\Theta$ is discussed in 
\cite{Friedrich2012}.

The incident beam is simple enough to define given our set up 
parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
\textit{Ott} uses a point matching approach to approximate the beam shape 
coefficients of the incident field by fitting it to the far field estimate. 
From the QPD's perspective it is receiving light both from the incident 
and scattered beam simultaneously, as such both fields must be expressed
using outgoing vector spherical harmonics.
\begin{align}
	E_{\rm inc}(r)&=
	E_0\sum^\infty_{n}\sum^n_{m=-n}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
	\\
	E_{\rm scat}(r)&=E_0 \sum^\infty_n \sum^n_{m=-n} \left[
	p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
	\right] 
\end{align}

Where the superscript (2) denotes an outgoing spherical harmonic
function. In order to compute the scattering from the target particle \textit{ott} uses the $T$-matrix method, this is not essential 
for a simple sphere but is essential for complex shaped particles 
such as dimers. The scattered and incident fields are then combined 
together in the far field to get $I(x,y)$, the quadrant and overall 
signals are calculated via:
\begin{align}
	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
\end{align}

Where the denominator is the total intensity on the QPD while there is 
no particle within the trap. You would expect that $S_x$ and $S_y$ are
near identical for equal particle displacements, however due to the 
polarisation of the beam there will be a slight bias to the signals 
for motion along the direction of the polarisation vector. This is 
generally not an issue if the particle trajectory is sampled for a 
long enough time period. By converting from signal units to length
units (see \eqref{eq:conversion_factor}) the trap shape can be 
discerned from the QPD signals, assuming the fluid properties are
known to a high degree of accuracy.

While translational motion has no knock on effects to the QPD signal, 
rotational changes are double counted by both the incident and total 
field. This results in rotational motion being biased in the QPD signal 
- even when collecting signals from isotropic scatterers. To prevent 
this we need to rotate the total field via the inverse rotation matrix
of the dimer.

>>>>

\chapter{Theory and methods}
\label{chapter:theory}
\textit{\textbf{Add overview here}} \\
This chapter covers the theory behind light and particle interactions, 
and how the exact theory varies for differently sized particles. After 
that we show how this can be applied to existing diffusion models to 
accurately simulate the behaviour of a particle within the vicinity of 
an optical trap. Lastly, we highlight the techniques utilised in 
experiments to both calibrate and characterise an optical trap based 
on our understanding of the trap.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Electromagnetism and optical tweezing}

Proper understanding of optical tweezing requires a description
of how the trapped particle interacts with the laser. From an
electromagnetism perspective the laser creates a spatially and 
temporally coherent electric field that scatters light when 
passing though a medium. The laws governing electric and magnetic 
fields are summarised most succinctly via the Maxwell equations. 
These four equations describe how the electric and magnetic fields 
behave and relate to one another. Any discussion of optical trapping 
is underpinned by the fact that in all cases the Maxwell equations 
must be satisfied. Macroscopic forms of the Maxwell equations are 
given below
\cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
where $\rho_f$ and $\mathbf{J}_f$ are the charge density and 
current caused by the presence of free charges in the medium, 
and $\mathbf{D}$ and $\mathbf{H}$ are the displacement and 
magnetising fields respectively. The latter two are defined by:
\begin{align}
	\mathbf{D} &= \epsilon\mathbf{E}+\mathbf{P} 
	\\ 
	\mathbf{H} &= \frac{1}{\mu}\mathbf{B}-\mathbf{M}
\end{align}
where $\mu$ is the vacuum permeability,, $\epsilon$ is the 
permittivity of free space, $\mathbf{P}$ is the polarization 
field and $\mathbf{M}$ is the magnetisation field, these 
fields arise due to the bound charges throughout the medium 
interacting with the EM field.

@@

The Lorenz-Mie theory provides an exact solution to the Maxwell 
equations for the scattering caused by an isotropic sphere. This 
theory describes the scattered wave given off by a dielectric 
sphere when incident by a plane wave as a summation of partial 
spherical waves. For any wave equation the vector fields must 
solve the Helmholtz equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align} 
where $k$ is the wave number of the electromagnetic radiation 
($k = \omega^2 \mu \epsilon/c^2$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a approximate Gaussian beam 
interacting with a sphere of radius $a$, the incident, internal, 
and scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r})
\end{align}

where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and ${\bf M}_
{nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
spherical harmonics respectively. The superscript denote whether 
the field is inward [(1)] or outward propagating [(3)]. By linearity 
of Maxwell's equations, its possible to relate the expansion 
coefficients of the scattered field directly to those of the incident 
field. As such we can write the scattering coefficients in terms of 
their incident components and the target's 'T-matrix'.
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} & T_{12} \\
		T_{21} & T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where the matrix \textbf{T} accounts for the targets shape, 
and refractive index. The $T$-matrix method can be easily 
applied to any solution of Maxwell's equations. In order to compute 
the optical force and torque applied to a given object one simply 
needs to know the momentum carried by the incident and scattered 
fields. So long as we know the expansion coefficients of the 
incident beam we can find the scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to compute the force.

@@

\noindent
Where $a_{nm}, b_{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size.
Though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are easier to approximate by alternative theories. When the 
particle radius is $\gg \lambda$ the optical force is best described 
by the Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (EBCM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the EBCM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix. Because the T-matrix is based only 
on the properties of the scatterer it can be reused in 
calculations regardless of the incident field in question.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker 
\cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh approximation, though in order to compute 
the optical force one now requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target. 
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Ray-Optics Regime}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rayleigh Scattering}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Describing the trajectory of a trapped particle}
While describing the optical forces is sufficient for 
theoretical characterisation of an optical trap, in an 
experimental setting the observed motion is not as 
simple. As such it is often far simpler to 
characterise the optical trap by recording the particle's
trajectory (see \ref{sec:position_detection} for 
specifics regarding the equipment), and then fitting it 
to our presumed description of the beam. The trajectory 
is easily understood if we understand the particle's 
diffusive behaviour.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of spherical aggregates}
\label{sec:simulations}
Later chapters cover the dynamics of spherical aggregates and anisotropic 
scatterers, these subjects are particularly difficult to characterise 
using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
As an example consider a symmetric dimer as a paradigmatic aggregate; 
if we consider the Langevin equation for such a aggregate within an optical 
trap we have:
\begin{align}
	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
	\sqrt{2\vec{D}_x}\eta(t)
\end{align}

Vigilante et al compiled together a python package that combines 
both \textit{ott} and \textit{mstm} to simulate the behaviour of 
spherical aggregates within a predefined optical trap 
\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
modules that can be used to model the scattering from different 
shaped particles; however, for aggregates of spheres the resultant 
force is nonsensical (see Chapter 4 for further details). As such 
we rely on the \textit{mstm} for simulating dimers.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Calibration Techniques}
\label{sec:calibration_techniques}
There are several approaches for calibrating and characterizing 
the optical trap, each approach has its drawbacks and benefits 
so each option should be chosen based on what elements want to 
be characterized. The basis for each of these methods stems from 
the analytical solution of the Langevin equation:
\begin{align}
	\label{eq:anylitical_lang}
	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
\end{align}

@@
